An investigation of the clinical utility of preoperative cardiopulmonary exercise testing in patients undergoing major pancreatic surgery.

### 2016-02-13 Presentations/viva_presentation_short.tex
Viva Presentation - First Draft Complete
- Email this to Donny
- Take a break
- Shower and clear toxins

@@

\begin{frame}
	\frametitle{Pancreatic cancer} %Epidemiology/outcomes
	\begin{block}{The Problem}
		\begin{itemize}
			\item 10th common but 5th common cause of cancer death
			\item Survival	1-yr: 20.8\%, 5-yr: 3.3\%, 10yr: 1.1\%
			\item Operable disease : approx 15\%
		\end{itemize}
	\end{block}
	\begin{block}{Treatment}
		\begin{itemize}
			\item Surgery + adjuvant chemotherapy only chance of cure
		\end{itemize}
	\end{block}
	\begin{block}{Pancreaticoduodenectomy}
		\begin{itemize}
			\item Complex surgery with 50\% morbidity
			\item Performed only in specialist centres
			\item Patient fitness is as important as tumour factors
		\end{itemize}
	\end{block}
\end{frame}

\begin{frame}
	\frametitle{Pancreaticoduodenectomy}
	\begin{columns}[t]
		\column{0.45\textwidth}
		\begin{figure}
			\centering
			\includegraphics[width=0.8\linewidth]{whipple_schematic_pre}
			\label{fig:whipple_schematic_pre}
		\end{figure}
		{\scriptsize
			\textbf{Structures removed:}
			\begin{itemize}
			\item Head of pancreas
			\item Distal stomach, duodenum, proximal jejunum
			\item  Common bile duct, gallbladder
			\item Lymphadenectomy
			\item Venous resection in some
		\end{itemize}}

		\column{0.45\textwidth}
		\begin{figure}
			\centering
			\includegraphics[width=0.8\linewidth]{whipple_schematic_post}
			\label{fig:whipple_schematic_post}
		\end{figure}
		{\scriptsize
		\textbf{Reconstruction:}
		\begin{itemize}
			\item Pancreaticojejunostomy (25\% leak rate)
			\item Hepaticojejunostomy
			\item Gastrojejunostomy
		\end{itemize}
		}
	\end{columns}
\end{frame}

\begin{frame}
	\frametitle{Complications of Surgery}
	\framesubtitle{Section 1.2.4 Pages 8-11}
	\begin{itemize}
		\item Significant improvements in mortality ($<$5\% from $>$40\%)
		\item Morbidity remains high at over 40\%
		\item Postoperative pancreatic fistula (POPF) is serious
		\item POPF can lead to other complications: Sepsis, PPH, DGE
		\item International Study Group for Pancreatic Surgery
		\item Clavien-Dindo classification of complications
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Adjuvant therapy}
	\framesubtitle{Section 1.3 Pages 11,13 }
	\begin{itemize}
		\item Adjuvant therapy improves survival
		\item Currently standard of care and subject of multiple trials
		\item 25\% - 50\% of patients do not receive adjuvant treatment
		\item Multiple factors including:
		\begin{itemize}
			\item Comorbidity
			\item Complications of surgery
			\item Favourable pathological features
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Risk Stratification }
	\framesubtitle{Section 1.4 Pages 13-18 }
	\begin{itemize}
		\item Patient fitness is important determinant of outcome
		\item Effect of comorbidity on function is difficult to assess
		\item POSSUM, APACHE, mGPS, E-PASS, etc.
		\item Static measures do not predict performance under stress
		\item Do not provide targets for prehabilitation
		\item Pancreas specific risk factors of POPF
		\begin{itemize}
			\item Soft pancreas
			\item Narrow pancreatic duct
			\item Pathology not obstructing PD
			\item These factors are not modifiable
			\item Coronary artery disease! (Lin 2004)
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{What is CPET?}
	\framesubtitle{Section 1.5 Page 18}
	\begin{itemize}
		\item Objective measure of cardiopulmonary response to exercise
		\item Dynamic test measuring numerous parameters
		\item Cardiac, respiratory, vascular and mitochondrial function
		\item First described for preoperative assessment by Older and co-workers
		\item Widely used in general, oesophago-gastric, vascular and thoracic surgery
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{CPET Methodology}
	\framesubtitle{Section 1.5.2 Page 20 }
	\begin{itemize}
		\item Dynamic exercise test
		\item Cycle ergometer with electronic braking and incremental workload
		\item Breath-by-breath gas analysis using a tight-fitting face mask
		\item Continuous monitoring of 12-lead ECG
		\item Initial 3-minute rest period on bike for baseline parameters
		\item Incremental work-load test to volitional fatigue
		%include picture here
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Physiological basis of CPET} 
	\begin{figure}
		\centering
		\includegraphics[width=0.5\linewidth]{CellRespiration}
		\label{fig:CellRespiration}
	\end{figure}
	\begin{block}
	\centering
	$H^+ + HCO3^- \Longleftrightarrow H_2CO_3 \Longleftrightarrow H_2O + CO_2$
	\end{block}

@@

\begin{frame}
	\frametitle{Determination of Anaerobic Threshold}
	\framesubtitle{Section 1.5.3 Page 22}
	\begin{figure}[htbp]
		\centering
		\includegraphics[width=0.7\linewidth]{../Figures/cpet_vslope}
		%\caption{Determination of $\dot{V}_{O_2}$AT by the V-slope method.}
		\label{fig:cpet_vslope}
	\end{figure}
\end{frame}

\begin{frame}
	\frametitle{Role of CPET in preoperative assessment}
	\begin{itemize}
		\item General Surgery
		\item Oesophago-gastric surgery
		\item Bariatric surgery
		\item Liver transplantation
		\item Thoracic surgery
		\medskip
		\item Little evidence in pancreatic surgery in 2010

	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{CPET in the Glasgow Cohort}
	\framesubtitle{Why was 'aerobic capacity' so poor in these patients yet outcomes were comparable?}
	\begin{figure}
		\centering
		\includegraphics[height=0.35\textheight]{../Figures/cpet_outcomes_dist_of_AT}
	\end{figure}

@@

\begin{frame}
\Huge{\centerline{The End}}
\end{frame}

>>>>

\begin{frame}
	\frametitle{My Supervisors - Thank you!}
	I wish to thank the following people for their unwavering support through the past few years. I thank them especially for their faith that I will see this through and if they had any doubts, for not telling me about them.\\
	\bigskip
	Professor Donald C McMillan,\\
	University Department of Surgery, Glasgow Royal Infirmary\\
	\bigskip
	Professor Paul Horgan,\\
	University Department of Surgery, Glasgow Royal Infirmary\\
\end{frame}

@@

\begin{frame}
	\frametitle{Pancreatic cancer} %Epidemiology/outcomes
	\begin{block}{The Problem}
		\begin{itemize}
			\item 10th common but 5th common cause of cancer death
			\item Survival	1-yr: 20.8\%, 5-yr: 3.3\%, 10yr: 1.1\%
			\item Operable disease : approx 15\%
		\end{itemize}
	\end{block}
	\begin{block}{Treatment}
		\begin{itemize}
			\item Surgery + adjuvant chemotherapy only chance of cure
		\end{itemize}
	\end{block}
	\begin{block}{Pancreaticoduodenectomy}
		\begin{itemize}
			\item Complex surgery with 50\% morbidity
			\item Performed only in specialist centres
			\item Patient fitness is as important as tumour factors
			\item Adjuvant treatment improves survival
		\end{itemize}
	\end{block}
\end{frame}

\begin{frame}
	\frametitle{Pancreaticoduodenectomy}
	\framesubtitle{Postoperative complications}
	\begin{itemize}
		\item Significant improvements in mortality ($<$5\% from $>$40\%)
		\item Morbidity remains high at over 40\%
		\item Postoperative pancreatic fistula (POPF) is serious
		\item POPF can lead to other complications: Sepsis, PPH, DGE
		\item International Study Group for Pancreatic Surgery
		\item Clavien-Dindo classification of complications
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Risk Stratification }
	\framesubtitle{Section 1.4 Pages 13-18 }
	\begin{itemize}
		\item Patient fitness is important determinant of outcome
		\item Effect of comorbidity on function is difficult to assess
		\item POSSUM, APACHE, mGPS, E-PASS, etc.
		\item Static measures do not predict performance under stress
		\item Do not provide targets for prehabilitation
		\item Preoperative management of obstructive jaundice remains controversial
		\item Pancreas specific risk factors of POPF
		\begin{itemize}
			\item Soft pancreas, narrow pancreatic duct, non-obstructing pathology
			\item These factors are not modifiable
			\item Coronary artery disease! (Lin 2004)
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Cardiopulmonary exercise testing}
	\framesubtitle{What, how, why?}
	\begin{itemize}

		\item Objective measure of cardiopulmonary response to exercise
		\item Dynamic test measuring multiple aspects of oxygen delivery
		\item Cardiac, respiratory, vascular and mitochondrial function
		\item First described for preoperative assessment by Older and co-workers
		\item Widely used in general, oesophago-gastric, vascular and thoracic surgery
		\item CPET Methodology
		\begin{itemize}
			\item Cycle ergometer with electronic braking
			\item Breath-by-breath gas analysis using a tight-fitting face mask
			\item Continuous monitoring of 12-lead ECG
			\item Initial 3-minute rest period on bike for baseline parameters
			\item Incremental work-load test to volitional fatigue
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Physiological basis of CPET} 
	\begin{columns}
	\column[c]{0.5\textwidth}

		\begin{figure}
			\includegraphics[width=\linewidth]{CellRespiration}
		\end{figure}
		{\scriptsize $H^+ + HCO3^- \Longleftrightarrow H_2CO_3 \Longleftrightarrow H_2O + CO_2$}

	\column[c]{0.5\textwidth}
	\begin{figure}[htbp]
		\includegraphics[width=\linewidth]{../Figures/cpet_vslope}
	\end{figure}

	\end{columns}

@@

@@

\begin{frame}
	\frametitle{CPET in the Glasgow Cohort}
	\framesubtitle{Why was 'aerobic capacity' so poor in these patients, yet outcomes were comparable?}
	\begin{figure}
		\centering
		\includegraphics[height=0.35\textheight]{../Figures/cpet_outcomes_dist_of_AT}
	\end{figure}

@@

\section{The End}

\begin{frame}
\frametitle{The End}
\framesubtitle{Tools and Roles}

\begin{table}
	\begin{tabular}{r l}
		        database & MS Access with VBA            \\
		data collection: & VBA + Dynamic Data Exchange   \\
		     statistics: & SPSS versions 19, 22          \\
		   bibliography: & Zotero, Bibus                 \\
		 thesis writing: & LaTeX using TeXstudio         \\
		version control: & Git + Dropbox                 \\
		                 &  \\
		         salary: & Surgical Critical Care Fellow \\
		                 & Surgical SpR on-call at GRI   \\
		                 & \\
		         sanity: & Staff on Surgical HDU, Donny
	\end{tabular}
\end{table}
